Improving real-time trend estimates using local parametrisation of polynomial regression filters

### 2024-04-30 paper/article_jos.qmd
v1

linestretch: 2
echo: false
tbl-cap-location: top
fig-cap-location: top
bibliography: biblio.bib
csl: chicago-author-date.csl
---

@@

:::: {.abstract}
This paper studies and compares real-time estimates of the trend-cycle component with moving averages based on local polynomial regression.
It enables to reproduce Henderson's symmetric and Musgrave's asymmetric filters used in the X-13ARIMA-SEATS seasonal adjustment algorithm.
This paper proposes two extensions of local polynomial filters for real-time trend-cycle estimates: first including a timeliness criterion to minimise the phase shift;
second, with procedure for parametrising asymmetric filters locally while they are generally parametrised globally, which can be suboptimal around turning points.
An empirical comparison, based on simulated and real data, shows that modelling polynomial trends that are too complex introduces more revisions without reducing the phase shift, and that local parametrisation reduces the delay in detecting turning points and reduces revisions.
The results are reproducible and all the methods can be easily applied using the R package `rjd3filters`.

@@

Analysis of the economic cycle, and in particular the early detection of turning points, is a major topic in the analysis of economic outlook.
To achieve this, economic indicators are generally seasonally adjusted.
However, in order to improve their interpretability, it may be necessary to perform additional smoothing to reduce noise, thereby to analyse the trend-cycle component.
By construction, trend-cycle extraction methods are closely related to seasonal adjustment methods, since they are generally applied to seasonally adjusted series. 
Thus, even though trend-cycle extraction methods are generally applied to seasonally adjusted series, the estimation of these series also depends on trend-cycle estimation methods.

A moving average, or linear filter, is a statistical method that consists of applying a rolling weighted mean to a times series: for each date $t$ it computes a weighted mean of $p$ past points and $q$ future points where $p,q\geq0$ depends on the moving average.
They are ubiquitous in business cycle extraction and seasonal adjustment methods.
For example, the X-11 seasonal adjustment method [@ladiray2011seasonal], used in X-13ARIMA-SEATS software [@x13], uses several moving averages to estimate the main components of a time series. 
Symmetric filters are applied to the centre of the series and for trend-cycle extraction, the most popular symmetric filter is the @henderson1916note moving average, which is used in the X-11 seasonal adjustment algorithm.

However, for real-time estimates, due to the lack of future observations, all these methods must rely on asymmetric filters to estimate the most recent points. 
For example, for trend-cycle extraction, X-11 uses Henderson's symmetric filter and @musgrave1964set's asymmetric filters on an extended series using an ARIMA model.
Since the predicted values are linear combinations of past values, this amounts to applying asymmetric moving averages at the end of the series.

If the classic asymmetric moving averages have good properties regarding the future revisions induced by the process (see for example @pierce1980SA), they create, by construction, phase shifts that impact the real-time estimation of turning points, introducing time delay in the detection.

Using local polynomial regression, @proietti2008 developed a general class of symmetric and asymmetric moving averages.
In particular, this class encompass Henderson's symmetric filter and Musgrave's asymmetric filters.
The aim of this study is to propose two extensions to the @proietti2008 class of filters in order to take account of two drawbacks of those methods.
Firstly,  adding, in the optimisation process, a timeliness criterion to directly control the phase shift (i.e., the delay in detecting turning points).
Secondly, calibrating asymmetric filter locally, while they are generally calibrated globally using thus assumptions that may be wrong locally.
Those extensions are implemented in the statistical software R [@logicielr] throw the `rjd3filters` package [@rjd3filters] and can therefore be easily used.
All the code associated to this paper is available at <https://github.com/XXXX> (anonymized linked before the manuscript is accepted).

@@

The basic assumption used in  time series decomposition methods is that the input observed time series, $y_t$ can be decomposed into a signal component $\mu_t$ and an erratic component $\varepsilon_t$ (called irregular):
$$
f(y_t)=\mu_t+\varepsilon_t
$$
where $f$ denotes some appropriate transformation (usually logarithmic or no transformation).
To simplify further notations, $y_t$ will denote the transformed observed series.
The noise component $\varepsilon_t$ is usually assumed to be a white noise.
Assuming that the input time series is seasonally adjusted (or without seasonality), the signal $\mu_t$ represents the trend and cycle components, here estimated jointly and called trend-cycle $TC_t$.
This component can be locally approximated by a polynomial degree (see @sec-nonparamreg):
$$
\mu_t = \sum_{j=0}^d\beta_jt^j+\xi_t
$$
with $\xi_t$ a stochastic process uncorrelated with $\varepsilon_t$.
Even if some papers model $\xi_t$ and $\varepsilon_t$ separately [see for example @GrayThomson2002], a usual assumption, adopted in this paper, is to is to gather $\xi_t$ and $\varepsilon_t$.
Thus, the trend-cycle $TC_t$ is considered as deterministic and modelled as a polynomial trend of degree $d$.
The coefficients $(\beta_0,\dots,\beta_k)$ can be estimated by weighted least squares.
The estimate $\hat \beta_0$ provides the trend-cycle estimate $\hat TC_t$ and it can be shown that this equivalent to apply of moving average (see @sec-nonparamreg).
The goal of this paper is to present two new methods to separate the trend-cycle component from the noise component in real-time (i.e., when no future point is known), using moving averages build on local polynomial smoothing techniques.

@@

Let $p$ and $f$ two integers, a moving average $M_{\boldsymbol\theta}$ or $M$ is defined by a set of coefficients $\boldsymbol\theta=\transp{\begin{pmatrix}\theta_{-p}&\cdots&\theta_{f}\end{pmatrix}}$ such as for all time series $y_t$:
$$
M_{\boldsymbol\theta}(y_t)=\sum_{k=-p}^{+f}\theta_kX_{t+k}.
$$

@@

To interpret the notions of gain and phase shift, it is useful to illustrate the effects of moving averages on harmonic series $y_t=\e^{-i\omega t}$ with $\omega\in[0,\pi]$.
The moving average $M_{\boldsymbol\theta}$ transforms $y_t$ into:
$$
Y_t = M_{\boldsymbol\theta}y_t = \sum_{k=-p}^{+f} \theta_k \e^{-i \omega (t+k)}
= \left(\sum_{k=-p}^{+f} \theta_k \e^{-i \omega k}\right)\cdot y_t.
$$
The function $\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{-i \omega k}$ is called the *transfer function* or *frequency response function*.
The frequency response function can alternatively be defined as $\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{i \omega k}$ or $\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{2\pi i \omega k}$.
It can be rewritten as:
$$
\Gamma_{\boldsymbol\theta}(\omega) = \rho_{\boldsymbol\theta}(\omega)\e^{i\varphi_{\boldsymbol\theta}(\omega)},
$$
where $\rho_{\boldsymbol\theta}(\omega)=G_{\boldsymbol\theta}(\omega)=\lvert\Gamma_{\boldsymbol\theta}(\omega)\rvert$ is the *gain* or *amplitude* function and $\varphi_{\boldsymbol\theta}(\omega)$ is the *phase shift* or *time shift* function.
We this have:
\begin{align*}
\rho_{\boldsymbol\theta}(\omega) &=\sqrt{\left(\sum_{k=-p}^{+f}\theta_k\cos(\omega k)\right)^2+
\left(\sum_{k=-p}^{+f}\theta_k\sin(\omega k)\right)^2} \\
\tan \varphi_{\boldsymbol\theta}(\omega) &=- \frac{\sum_{k=-p}^{+f}\theta_k\sin(\omega k)}{\sum_{k=-p}^{+f}\theta_k\cos(\omega k)} \text{ if }\sum_{k=-p}^{+f}\theta_k\cos(\omega k) \ne 0 \\
&= 0 \text{ if }\sum_{k=-p}^{+f}\theta_k\cos(\omega k) = 0.
\end{align*}
The phase shift is sometimes represented as $\phi_{\boldsymbol\theta}(\omega)=\frac{\varphi_{\boldsymbol\theta}(\omega)}{\omega}$ to measure it in number of periods.
For all symmetric moving averages we have $\varphi_{\boldsymbol\theta}(\omega)\equiv 0 \;(modulo\;{\pi})$.

@@

Fourier decomposition allows us to analyze any time series as a sum of harmonic series, and each component (trend, cycle, seasonal, irregular) is associated with a set of frequencies.
Noting $\omega = 2\pi/s$, the harmonic series of frequency $\omega$ represents a series that repeats itself every $s$ periods.
For example, for a monthly series ($s=12$ observations per year), seasonal movements are those that repeat every year: they are therefore associated with frequencies $2\pi/12$ (annual periodicity), $2\pi/12\times 2=2\pi/6,\dots,2\pi/12\times 5$.
In this paper, we consider that the trend-cycle is associated with frequencies in the interval $[0, 2\pi/12]$, i.e., movements recurring at least every 12 months.
Even if different frequencies are sometimes used (e.g. $[0, 2\pi/36]$ to consider only cycles of at least 36 months), this has no impact on the results of the study.
The other frequencies, $[2\pi/12, \pi]$ are associated with the irregular component.

@@

\justifying
*Lecture: on monthly times series, 12-month cycles (associated to the frequency $2\pi/12=\pi/6$) are amplified of 10% ($G_{\boldsymbol\theta}(\pi/6)=1.1$) and associated turning points are detected with a delay of 1 months ($\varphi_{\boldsymbol\theta}(\pi/6)/(\pi/6)\simeq -1$).*\
*Note: the Musgrave filter depends on a ratio $I/C$ (see @sec-nonparamreg).* 
*In X-11, for monthly series, when the bandwidth is 6 (i.e.: a symmetric Henderson filter of order 13 is used, what is used in most cases), the ratio is fixed to 3.5 and this value is used in this figure.*

Coefficients, gain and phase shift function for the Musgrave filter used in X-11 in real time (last estimate) for monthly series when the bandwidth is 6.
:::

@@

Trend-cycles are generally modelled by local polynomial trends (see @sec-nonparamreg) and, in order to best preserve the trend-cycles, we thus design moving averages that preserve the polynomial trends.
A moving average $M_{\boldsymbol\theta}$ preserves a function of time $f(t)$ if $\forall t:M_{\boldsymbol\theta} f(t)=f(t)$.

For the moving average $M_{\boldsymbol\theta}$, it can be shown that:

- $M_{\boldsymbol\theta}$ preserves polynomials of degree $d$ if and only if:
$$
\sum_{k=-p}^{+f}\theta_k=1
 \text{ and }
\forall j \in \left\{1,2,\dots,d\right\}:\:
\sum_{k=-p}^{+f}k^j\theta_k=0.
$$
- If $M_{\boldsymbol\theta}$ is symmetric ($p=f$ and $\theta_{-k} = \theta_k$) and preserves polynomials of degree $2d$ then it also preserves polynomials of degree $2d+1$.

@@

2. Apply symmetric filters to extended forecast series.
This method seems to date back to @deforest1877adjustment, which also suggests modelling a polynomial trend of degree three or less at the end of the period.
A similar approach is used in the X-13ARIMA-SEATS seasonal adjustment method: the series is extended by one year (by default) with an ARIMA model to minimise the revisions associated to asymmetric filters.
Since the final filter used for the seasonal adjustment extraction is much longer than one year (in most of the cases, for monthly series, the bandwidth is 78), the input series is just partially extended (and asymmetric filters are still used) to prevent from instability over long periods (e.g., outliers or model shifts).

Ultimately, the second method is equivalent to using asymmetric moving averages, since forecasts are linear combinations of the past.

Conversely, the implicit forecasts of an asymmetric moving average can be deduced from a reference symmetric moving average. 
This allows us to judge the quality of real-time estimates of the trend-cycle and to anticipate future revisions when forecasts are far from expected evolutions.\
Let us denote $\boldsymbol v=(v_{-h},\dots, v_{h})$ the reference symmetric moving average (used to compute the final estimates) and $\boldsymbol w^0,\dots, \boldsymbol w^{h-1}$ a sequence of asymmetric moving averages used for the intermediate estimates. 
$\boldsymbol w^0=(w_{-h}^0, \dots, w_{0}^0)$ is used for estimation in real time estimate (when 0 point in the future is known), $\boldsymbol w^1=(w_{-h}^1, \dots, w_{0}^1, w_{1}^1)$ is used for estimation of the penultimate point (when only 1 in the future is known), and so on.
Let $y_{t-h},\dots,y_{t}$ be the observed series during the last $h$ periods (for which asymmetric moving averages are used to get the intermediate trend-cycle estimates for the dates $t-h$ to $h$). 
The implicit forecast $y_{t+1}^*,\dots y_{t+h}^*$ induced by $w^0,\dots w^{h-1}$ are the forecasts of the series $y_t$ for which the estimates of the symmetric filter applied to the extended series $\boldsymbol y^* = (y_{t-h}, \dots, y_t, y_{t+1}^*,\dots y_{t+h}^*)$ produce the same estimates as the asymmetric filter applied to the series $\boldsymbol y^*$.
This means:
\begin{align*}
\underbrace{\sum_{i=-h}^0 v_iy_{ti} + \sum_{i=1}^h v_iy_{t+i}^*}_{\text{smoothing by }\boldsymbol v\text{ of the extended series}} &=
\underbrace{\sum_{i=-h}^0 w_i^0y_{t+i}}_{\text{smoothing by }\boldsymbol w^0\text{ using }0\text{ points after }t} \\
&=
\underbrace{\sum_{i=-h}^0 w_i^1y_{t+i} + w_1^1y_{t+1}^*}_{\text{smoothing by }\boldsymbol w^1\text{ using }1\text{ points after }t} \\
&= \underbrace{\sum_{i=-h}^0 w_i^2y_{t+i} + \sum_{i=1}{2}w_i^2y_{t+i}^*}_{\text{smoothing by }\boldsymbol w^2\text{ using }2\text{ points after }t} \\
&= \cdots \\
&=\underbrace{\sum_{i=-h}^0 w_i^{h-1}y_{t+i} + \sum_{i=1}^{h-1} w_i^{h-1}y_{t+i}^*}_{\text{smoothing by }\boldsymbol w^{h-1}\text{ using }h-1\text{ points after }t}
\end{align*}
For any integer $q$ between 0 and $h-1$, using for convention $w_t^q=0$ if $t>q$ we then have:
$$
\sum_{i=1}^h (v_i- w_i^q) y_i^*
=\sum_{i=-h}^0 (w_i^q-v_i)y_i.
$$
In matrix form, this is equivalent to solving:
$$
\scriptstyle
\begin{pmatrix}
  v_1 & v_2 & \cdots & v_h \\
  v_1 - w_1^1 & v_2 & \cdots & v_h \\
  \vdots & \vdots & \cdots & \vdots \\
   v_1 - w_1^{h-1} & v_2-w_2^{h-1} & \cdots & v_h
\end{pmatrix}
\begin{pmatrix}y_1^* \\ \vdots \\ y_h^*\end{pmatrix}=
\begin{pmatrix}
  w_{-h}^0 - v_{-h} & w_{-(h-1)}^0 - v_{-(h-1)} & \cdots & w_{0}^0 - v_{0} \\
  w_{-h}^1 - v_{-h} & w_{-(h-1)}^1 - v_{-(h-1)} & \cdots & w_{0}^1 - v_{0} \\
  \vdots & \vdots & \cdots & \vdots \\
  w_{-h}^{h-1} - v_{-h} & w_{-(h-1)}^{h-1} - v_{-(h-1)} & \cdots & w_{0}^{h-1} - v_{0}
\end{pmatrix}
\begin{pmatrix}y_{-h} \\ \vdots \\ y_0\end{pmatrix}.
$$
This is implemented in the `rjd3filters::implicit_forecast()` function.

@@

Many trend-cycle extraction methods are based on non-parametric regressions, which are particularly flexible because they do not assume any predetermined dependency in the predictors. 
In practice, local regressions can be used. 
More specifically, consider a set of points $(x_t,y_t)_{1\leq t\leq n}$ and we here consider that $x_t=t$ (no external regressor is used). 
Non-parametric regression involves assuming that there exists a function $\mu$, to be estimated, such that $y_t=\mu(t)+\varepsilon_t$ with $\varepsilon_t$ an error term.
According to Taylor's theorem, for any date $t_0$, if $\mu$ is differentiable $d$ times, then:
$$
\forall t \::\:\underbrace{\mu(t)}_{=TC_t} = \underbrace{\mu(t_0)}_{=TC_{t_0}} + \mu'(t_0)(t-t_0)+\dots +
\frac{\mu^{(d)}(t_0)}{d!}(t-t_0)^d+R_d(t),
$$ {#eq-taylor}
where $R_d$ is a negligible residual term in the neighbourhood of $t_0$. 
In a neighbourhood $h(t_0)$ around $t_0$, $\mu$ can therefore be approximated by a polynomial of degree $d$. 
The quantity $h(t_0)$ is called the *bandwidth*.
If $\varepsilon_t$ is white noise, we can estimate $\mu(t_0)=TC_{t_0}$ by least squares using the observations which are in $\left[t_0-h(t_0),t_0+h(t_0)\right]$. 
In practice, this means assuming that the trend is locally polynomial.
We also generally assume that the bandwith is fixed over time: $h(t_0)=h.$
This is also what is assumed in the seasonal adjustment algorithm X-11 and what will be assumed in this paper.

Various estimation methods can be used to derive symmetric and asymmetric moving averages.  
@GrayThomson1996 propose a complete statistical framework that makes it possible, in particular, to model the error in approximating the trend by local polynomials.
However, as the specification of this error is generally complex, simpler models may be preferred, such as that of @proietti2008.  
Finally, @dagumbianconcini2008 propose a similar modelling of the trend-cycle but using the theory of Hilbert spaces with reproducing kernels for estimation, which has the particular advantage of facilitating the calculation of different moving averages at different time frequencies.
See @inseeDTM202401 for a more detailed comparison of the different recent methods for trend-cycle extraction included in the `rjd3filters` package

>>>>

lang: en-GB  
linestretch: 2
echo: false
tbl-cap-location: top
fig-cap-location: top
bibliography: biblio.bib
csl: chicago-author-date.csl
---

@@

:::: {.abstract}
This paper examines and compares real-time estimates of the trend-cycle component using moving averages constructed with local polynomial regression.
It enables to reproduce Henderson's symmetric and Musgrave's asymmetric filters used in the X-13ARIMA-SEATS seasonal adjustment algorithm.
This paper proposes two extensions of local polynomial filters for real-time trend-cycle estimates: first including a timeliness criterion to minimise the phase shift;
second with procedure for parametrising asymmetric filters locally while they are generally parametrised globally, which can be suboptimal around turning points.
An empirical comparison, based on simulated and real data, shows that modelling polynomial trends that are too complex introduces more revisions without reducing the phase shift, and that local parametrisation reduces the delay in detecting turning points and reduces revisions.
The results are reproducible and all the methods can be easily applied using the R package `rjd3filters`.

@@

Analysis of the economic cycle, and in particular the early detection of turning points, is a major topic in the analysis of economic outlook.
To achieve this, economic indicators are generally seasonally adjusted.
However, in order to improve their interpretability, it may be necessary to perform additional smoothing to reduce noise, thereby to analyse the trend-cycle component.
By construction, trend-cycle extraction methods are closely related to seasonal adjustment methods, since they are generally applied to seasonally adjusted series.

A moving average, or linear filter, is a statistical method that consists of applying a rolling weighted mean to a times series: for each date $t$ it computes a weighted mean of $p$ past points and $q$ future points where $p,q\geq0$ depends on the moving average.
They are ubiquitous in business cycle extraction and seasonal adjustment methods.
For example, the X-11 seasonal adjustment method [@ladiray2011seasonal], used in X-13ARIMA-SEATS software [@x13], uses several moving averages to estimate the main components of a time series. 
Symmetric filters are applied to the centre of the series and for trend-cycle extraction, the most popular symmetric filter is the @henderson1916note moving average, which is used in the X-11 seasonal adjustment algorithm. 
However, for real-time estimates, due to the lack of future observations, all these methods must rely on asymmetric filters to estimate the most recent points. 
For example, for trend-cycle extraction, X-11 uses Henderson's symmetric filter and @musgrave1964set's asymmetric filters on an extended series using an ARIMA model.
Since the predicted values are linear combinations of past values, this amounts to applying asymmetric moving averages at the end of the series.
If the classic asymmetric moving averages have good properties regarding the future revisions induced by the process (for example @pierce1980SA shows similar performances in terms of revisions of X-11 compared to theoretically superior procedure based on ARIMA forecast models), they create, by construction, phase shifts that impact the real-time estimation of turning points, introducing time delay in the detection.

Using local polynomial regression, @proietti2008 developed a general class of symmetric and asymmetric moving averages.
In particular, this class encompass Henderson's symmetric filter and Musgrave's asymmetric filters.
The aim of this study is to propose two extensions to the @proietti2008 class of filters in order to take account of two drawbacks of those methods.
Firstly  adding, in the optimisation process, a timeliness criterion to directly control the phase shift (i.e., the delay in detecting turning points).
Secondly calibrating asymmetric filters locally, while they are generally calibrated globally using thus assumptions that may be wrong locally.
Those extensions are implemented in the statistical software R [@logicielr] throw the `rjd3filters` package [@rjd3filters] and can therefore be easily used.
All the code associated to this paper is available at <https://github.com/XXXX> (anonymized linked before the manuscript is accepted).

@@

The basic assumption used in  time series decomposition methods is that the input observed time series, $y_t$, can be decomposed into a signal component $\mu_t$ and an erratic component $\varepsilon_t$ (called irregular):
$$
f(y_t)=\mu_t+\varepsilon_t
$$
where $f$ denotes some appropriate transformation (usually logarithmic or no transformation).
To simplify further notations, $y_t$ will denote the transformed observed series.
The noise component $\varepsilon_t$ is usually assumed to be a white noise.
Assuming that the input time series is seasonally adjusted (or without seasonality), the signal $\mu_t$ represents the trend (sustained and systematic variations over a long period of time) and cycle (cyclical movements superimposed upon the long-term trend) components, here estimated jointly and called trend-cycle $TC_t$.
This component can be locally approximated by a polynomial degree (@sec-nonparamreg):
$$
\mu_t = \sum_{j=0}^d\beta_jt^j+\xi_t
$$
with $\xi_t$ a stochastic process uncorrelated with $\varepsilon_t$.
Even if some papers model $\xi_t$ and $\varepsilon_t$ separately [see for example @GrayThomson2002], a usual assumption, adopted in this paper, is to is to gather $\xi_t$ and $\varepsilon_t$.
Thus, the trend-cycle $TC_t$ is considered as deterministic and modelled as a polynomial trend of degree $d$.
The coefficients $(\beta_0,\dots,\beta_k)$ can be estimated by weighted least squares.
The estimate $\hat \beta_0$ provides the trend-cycle estimate $\widehat{TC}_t$ and it can be shown that this equivalent to apply of moving average (@sec-nonparamreg).
The goal of this paper is to present two new methods to separate the trend-cycle component from the noise component in real-time (i.e., when no future point is known), using moving averages build on local polynomial smoothing techniques.

@@

Let $p$ and $f$ two integers, a moving average $M_{\boldsymbol\theta}$ is defined by a set of coefficients $\boldsymbol\theta=\transp{\begin{pmatrix}\theta_{-p}&\cdots&\theta_{f}\end{pmatrix}}$ such as for all time series $y_t$:
$$
M_{\boldsymbol\theta}(y_t)=\sum_{k=-p}^{+f}\theta_kX_{t+k}.
$$

@@

To interpret the notions of gain and phase shift, it is useful to illustrate the effects of moving averages on harmonic series $y_t=\e^{-i\omega t}$ with $\omega\in[0,\pi]$.
The moving average $M_{\boldsymbol\theta}$ transforms $y_t$ into:
$$
Y_t = M_{\boldsymbol\theta}y_t = \sum_{k=-p}^{+f} \theta_k \e^{-i \omega (t+k)}
= \left(\sum_{k=-p}^{+f} \theta_k \e^{-i \omega k}\right)\cdot y_t.
$$
The function $\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{-i \omega k}$ is called the *transfer function* or *frequency response function*.
The frequency response function can alternatively be defined as $\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{i \omega k}$ or $\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{2\pi i \omega k}$.
It can be rewritten as:
$$
\Gamma_{\boldsymbol\theta}(\omega) = \rho_{\boldsymbol\theta}(\omega)\e^{i\varphi_{\boldsymbol\theta}(\omega)},
$$
where $\rho_{\boldsymbol\theta}(\omega)=G_{\boldsymbol\theta}(\omega)=\lvert\Gamma_{\boldsymbol\theta}(\omega)\rvert$ is the *gain* or *amplitude* function and $\varphi_{\boldsymbol\theta}(\omega)$ is the *phase shift* or *time shift* function.
We thus have:
\begin{align*}
\rho_{\boldsymbol\theta}(\omega) &=\sqrt{\left(\sum_{k=-p}^{+f}\theta_k\cos(\omega k)\right)^2+
\left(\sum_{k=-p}^{+f}\theta_k\sin(\omega k)\right)^2} \\
\tan \varphi_{\boldsymbol\theta}(\omega) &=- \frac{\sum_{k=-p}^{+f}\theta_k\sin(\omega k)}{\sum_{k=-p}^{+f}\theta_k\cos(\omega k)} \text{ if }\sum_{k=-p}^{+f}\theta_k\cos(\omega k) \ne 0 \\
&= 0 \text{ if }\sum_{k=-p}^{+f}\theta_k\cos(\omega k) = 0.
\end{align*}
The phase shift is sometimes represented as $\phi_{\boldsymbol\theta}(\omega)=\frac{\varphi_{\boldsymbol\theta}(\omega)}{\omega}$ to measure it in number of periods.
For all symmetric moving averages we have $\varphi_{\boldsymbol\theta}(\omega)\equiv 0 \;(modulo\;{\pi})$.

@@

Fourier decomposition allows us to analyse any time series as a sum of harmonic series, and each component (trend, cycle, seasonal, irregular) is associated with a set of frequencies.
With $\omega = 2\pi/s$, the harmonic series of frequency $\omega$ represents a series that repeats itself every $s$ periods.
For example, for a monthly series ($s=12$ observations per year), seasonal movements are those that repeat every year: they are therefore associated with frequencies $2\pi/12$ (annual periodicity), $2\pi/12\times 2=2\pi/6,\dots,2\pi/12\times 5$.
In this paper, we consider that the trend-cycle is associated with frequencies in the interval $[0, 2\pi/12]$, i.e., movements recurring at least every 12 months.
Even if different frequencies are sometimes used (e.g. $[0, 2\pi/36]$ to consider only cycles of at least 36 months), this has no impact on the results of the study.
The other frequencies, $[2\pi/12, \pi]$ are associated with the irregular component.

@@

\justifying
*Lecture: on monthly times series, 12-month cycles (associated to the frequency $2\pi/12=\pi/6$) are amplified by 10% ($G_{\boldsymbol\theta}(\pi/6)\simeq1.1$) and associated turning points are detected with a delay of 1 months ($\varphi_{\boldsymbol\theta}(\pi/6)/(\pi/6)\simeq -1$).*
*8-month cycles (associated to the irregular) are only reduced by 6% ($G_{\boldsymbol\theta}(\pi/4)\simeq0.94$).*\
*Note: the Musgrave filter depends on a ratio denoted $I/C$ (see @sec-nonparamreg).* 
*In X-11, for monthly series, when the bandwidth is 6 (i.e.: a symmetric Henderson filter of order 13 is used, what is used in most cases), the ratio is fixed to 3.5 and this value is used in this figure.*

Coefficients, gain and phase shift function for the Musgrave filter used in X-11 in real-time (last estimate) for monthly series when the bandwidth is 6.
:::

@@

Trend-cycles are generally modelled by local polynomial trends (see @sec-nonparamreg) and, in order to best preserve the trend-cycles, we thus design moving averages that preserve the polynomial trends.
A moving average $M_{\boldsymbol\theta}$ preserves a function of time $f(t)$ if $\forall t,M_{\boldsymbol\theta} f(t)=f(t)$.

For a moving average $M_{\boldsymbol\theta}$, it can be shown that:

- $M_{\boldsymbol\theta}$ preserves polynomials of degree $d$ if and only if:
$$
\sum_{k=-p}^{+f}\theta_k=1
 \text{(preserves constant) and }
\forall j \in \left\{1,2,\dots,d\right\},\:
\sum_{k=-p}^{+f}k^j\theta_k=0.
$$
- If $M_{\boldsymbol\theta}$ is symmetric ($p=f$ and $\theta_{-k} = \theta_k$) and preserves polynomials of degree $2d$ then it also preserves polynomials of degree $2d+1$.

@@

2. Apply symmetric filters to an extended forecast series (which is equivalent to using asymmetric moving averages, since forecasts are linear combinations of the past).
This method seems to date back to @deforest1877adjustment, which also suggests modelling a polynomial trend of degree three or less at the end of the period.
A similar approach is used in the X-13ARIMA-SEATS seasonal adjustment method: the series is extended by one year (by default) with an ARIMA model to minimise the revisions associated to asymmetric filters.
Since the final filter used for the seasonal adjustment extraction is much longer than one year (in most of the cases, for monthly series, the bandwidth is 78), the input series is just partially extended (and asymmetric filters are still used) to prevent from instability over long periods (e.g., outliers or model shifts).

Conversely, the *implicit forecasts* of an asymmetric moving average can be deduced from a reference symmetric moving average. 
This allows us to judge the quality of real-time estimates of the trend-cycle and to anticipate future revisions when forecasts are far from expected evolutions.\
Let us denote $\boldsymbol v=(v_{-h},\dots, v_{h})$ the reference symmetric moving average (used to compute the final estimates) and $\boldsymbol w^0,\dots, \boldsymbol w^{h-1}$ a sequence of asymmetric moving averages used for the intermediate estimates. 
$\boldsymbol w^0=(w_{-h}^0, \dots, w_{0}^0)$ is used for estimation in real-time estimate (when 0 point in the future is known), $\boldsymbol w^1=(w_{-h}^1, \dots, w_{0}^1, w_{1}^1)$ is used for estimation of the penultimate point (when only 1 in the future is known), and so on.
Let $y_{t-h},\dots,y_{t}$ be the observed series during the last $h$ periods (for which asymmetric moving averages are used to get the intermediate trend-cycle estimates for the dates $t-h$ to $t$). 
The *implicit forecast* $y_{t+1}^*,\dots y_{t+h}^*$ induced by $w^0,\dots w^{h-1}$ are the forecasts of the series $y_t$ for which the estimates of the symmetric filter applied to the extended series $\boldsymbol y^* = (y_{t-h}, \dots, y_t, y_{t+1}^*,\dots y_{t+h}^*)$ produce the same estimates as the asymmetric filter applied to the series $\boldsymbol y^*$.
This means:
\begin{align*}
\underbrace{\sum_{i=-h}^0 v_iy_{t+i} + \sum_{i=1}^h v_iy_{t+i}^*}_{\text{smoothing by }\boldsymbol v\text{ of the extended series}} &=
\underbrace{\sum_{i=-h}^0 w_i^0y_{t+i}}_{\text{smoothing by }\boldsymbol w^0\text{ using }0\text{ point after }t} \\
&=
\underbrace{\sum_{i=-h}^0 w_i^1y_{t+i} + w_1^1y_{t+1}^*}_{\text{smoothing by }\boldsymbol w^1\text{ using }1\text{ point after }t} \\
&= \underbrace{\sum_{i=-h}^0 w_i^2y_{t+i} + \sum_{i=1}^{2}w_i^2y_{t+i}^*}_{\text{smoothing by }\boldsymbol w^2\text{ using }2\text{ points after }t} \\
&= \cdots \\
&=\underbrace{\sum_{i=-h}^0 w_i^{h-1}y_{t+i} + \sum_{i=1}^{h-1} w_i^{h-1}y_{t+i}^*}_{\text{smoothing by }\boldsymbol w^{h-1}\text{ using }h-1\text{ points after }t}
\end{align*}
For any integer $q$ between 0 and $h-1$, using for convention $w_t^q=0$ if $t>q$ we then have:
$$
\sum_{i=1}^h (v_i- w_i^q) y_i^*
=\sum_{i=-h}^0 (w_i^q-v_i)y_i.
$$
In matrix form, this is equivalent to solving:
$$
\scriptstyle
\begin{pmatrix}
  v_1 & v_2 & \cdots & v_h \\
  v_1 - w_1^1 & v_2 & \cdots & v_h \\
  \vdots & \vdots & \cdots & \vdots \\
   v_1 - w_1^{h-1} & v_2-w_2^{h-1} & \cdots & v_h
\end{pmatrix}
\begin{pmatrix}y_1^* \\ \vdots \\ y_h^*\end{pmatrix}=
\begin{pmatrix}
  w_{-h}^0 - v_{-h} & w_{-(h-1)}^0 - v_{-(h-1)} & \cdots & w_{0}^0 - v_{0} \\
  w_{-h}^1 - v_{-h} & w_{-(h-1)}^1 - v_{-(h-1)} & \cdots & w_{0}^1 - v_{0} \\
  \vdots & \vdots & \cdots & \vdots \\
  w_{-h}^{h-1} - v_{-h} & w_{-(h-1)}^{h-1} - v_{-(h-1)} & \cdots & w_{0}^{h-1} - v_{0}
\end{pmatrix}
\begin{pmatrix}y_{-h} \\ \vdots \\ y_0\end{pmatrix}.
$$
This is implemented in the `rjd3filters::implicit_forecast()` function.

@@

Many trend-cycle extraction methods are based on non-parametric regressions, which are particularly flexible because they do not assume any predetermined dependency in the predictors. 
In practice, local regressions can be used. 
More specifically, consider a set of points $(x_t,y_t)_{1\leq t\leq n}$ and we here consider that $x_t=t$ (no external regressor is used). 
Non-parametric regression involves assuming that there exists a function $\mu$, to be estimated, such that $y_t=\mu(t)+\varepsilon_t$ with $\varepsilon_t$ an error term.
According to Taylor's theorem, for any date $t_0$, if $\mu$ is differentiable $d$ times, then:
$$
\forall t,\:\mu(t) = \underbrace{\mu(t_0)}_{=TC_{t_0}} + \mu'(t_0)(t-t_0)+\dots +
\frac{\mu^{(d)}(t_0)}{d!}(t-t_0)^d+R_d(t),
$$ {#eq-taylor}
where $R_d$ is a negligible residual term in the neighbourhood of $t_0$. 
In a neighbourhood $h(t_0)$ around $t_0$, $\mu$ can therefore be approximated by a polynomial of degree $d$. 
The quantity $h(t_0)$ is called the *bandwidth*.
If $\varepsilon_t$ is white noise, we can estimate $\mu(t_0)=TC_{t_0}$ by least squares using the observations which are in $\left[t_0-h(t_0),t_0+h(t_0)\right]$. 
In practice, this means assuming that the trend is locally polynomial.
We also generally assume that the bandwidth is fixed over time: $h(t_0)=h.$
This is also what is assumed in the seasonal adjustment algorithm X-11 and what will be assumed in this paper.

Various estimation methods can be used to derive symmetric and asymmetric moving averages.
For example, @GrayThomson1996 propose a complete statistical framework that makes it possible, in particular, to model the error in approximating the trend by local polynomials.
However, as the specification of this error is generally complex, simpler models may be preferred, such as that of @proietti2008.
@dagumbianconcini2008 propose a similar modelling of the trend-cycle but using the theory of Hilbert spaces with reproducing kernels for estimation, which has the particular advantage of facilitating the calculation of different moving averages at different time frequencies.
See @inseeDTM202401 for a more detailed comparison of the different recent methods for trend-cycle extraction included in the `rjd3filters` package.
